SMT-Boosted Security Types for Low-Level MPC

### 2025-11-07 toplas26/smt.tex
checkpoint

\subsubsection{Example: Correctness of 3-Party Addition}
Consider the following 3-party additive secret sharing protocol, where
each party generates two additive secret shares
\cite{evans2018pragmatic} for their respective secrets. Each party
publicly reveals the sums all their shares, and the sum of those is
the protocol output for each party.  
$$
\begin{array}{c@{\qquad}c}
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array} & 
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
\end{array}
$$
Letting $\prog$ be this protocol, we can verify correctness
as follows-- that is, the output sum of all shares is the sum of the
three input secrets:
$$
\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

@@

\begin{proof}
  Assume $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable without \TirName{Hyp}.
  Then $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$, since all the other rules
  are well-known to be validity preserving. 
  Now since $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$ by assumption, we have
  $
  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 
  $,
  and clearly:
  $$
  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge (\eqs_2 \impl \eqs)
  $$
  and since $\eqs_2 \wedge (\eqs_2 \impl \eqs)$ iff
  $\eqs_2 \wedge \eqs$, we have
  $\eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge \eqs$.
\end{proof}

>>>>

\subsubsection{Example: Correctness of 3-Party Addition}
\label{section-smt-example}
Consider the following 3-party additive secret sharing protocol, where
each party generates two additive secret shares
\cite{evans2018pragmatic} for their respective secrets. Each party
publicly reveals the sums all their shares, and the sum of those is
the protocol output for each party.  
$$
\begin{array}{c@{\qquad}c}
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
  \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
  \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
  \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
  \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
\end{array} & 
\begin{array}{lll}
  \rvl{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
  \rvl{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
  \rvl{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} \\
  \out{1} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{1}\\
  \out{2} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{2}\\
  \out{3} &:=& \elab{(\rvl{1} \fplus \rvl{2} + \rvl{3})}{3}
\end{array}
\end{array}
$$
Letting $\prog$ be this protocol, we can verify correctness
as follows-- that is, the output sum of all shares is the sum of the
three input secrets:
$$
\toeq{\prog}\ \models\ \out{1} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

@@

\begin{proof}
  \cnote{de-obfuscate.}
  Assume $\htrip{\eqs_1}{\prog}{\eqs_2}$ is derivable without \TirName{Hyp}.
  Then $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$, since all the other rules
  are well-known to be validity preserving. 
  Now since $\eqs_1 \wedge \toeq{\prog} \models \eqs_2$ by assumption, we have
  $
  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 
  $,
  and clearly:
  $$
  \eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge (\eqs_2 \impl \eqs)
  $$
  and since $\eqs_2 \wedge (\eqs_2 \impl \eqs)$ iff
  $\eqs_2 \wedge \eqs$, we have
  $\eqs_1 \wedge (\eqs_2 \impl \eqs) \wedge \toeq{\prog} \models \eqs_2 \wedge \eqs$.
\end{proof}

### 2025-11-07 toplas26/sty.tex
checkpoint

@@

$$
\begin{array}{rcl@{\hspace{10mm}}l}
  \sty &::=& \sharety \mid \pubty \mid \outty & \textit{share types} \\
  \stt &::=& [\phi,\phi] \mid \pubx{\phi} & \textit{share terms} 
\end{array}
$$

\subsection{Non-interactive typing rules}

\subsection{Interactive typing rules}

@@

@@

\begin{definition} Define $\ov(\Gamma)$ as follows:
\begin{eqnarray*}
\ov(\varnothing) &=& \varnothing \\[-1.5mm]
\ov(\setit{[\mx{z}{\cid}, \rx{y}{\cid'}] : \sharety} \cup \Gamma) &=& \setit{\mx{z}{\cid}} \cup \ov(\Gamma)\\[-1.5mm]
\ov(\setit{[\phi + \px{x}] : \pubty} \cup \Gamma) &=& \setit{\px{x}} \cup \ov(\Gamma)\\[-1.5mm]
\ov(\setit{[\phi] : \outty} \cup \Gamma) &=& \setit{\phi} \cup \ov(\Gamma)
\end{eqnarray*}
\end{definition}

>>>>

Our approach to static analysis is to enforce the use of sound
techniques in semi-homorphic encryption in protocols.  As discussed in
the example in Section \ref{section-smt-example}, protocols integrate
interactive elements, such as the sharing of one-time-pad (OTP)
encrypted secrets, and noninteractive elements, such as addition (of
shares). We approach this by introducing distinct judgement forms for
interactive and noninteractive protocol elements. Our type system
as presented will assume 2-client protocols, but this is for
simplicity and the analysis can be generalized to more parties.

Both interactive and non-interactive judgements leverage the use
of SMT interpretations as described in Section \ref{section-smt}.
We introduce a \emph{share type} language $\sty$, classifying
\emph{share terms}. We represent share terms using standard notation
\cite{10.1007/978-3-030-68869-1_3}. A share term $[\phi_1,\phi_2]$
represents a secret-shared value where client 1 has $\phi_1$ and
client 2 has $\phi_2$, and this shared value can be reconstructed
as $\phi_1 + \phi_2$.\cnote{Comment on generation and
  reconstruction in general somewhere.} A share term $\pubx{\phi}$
denotes a publicly shared value.
$$
\begin{array}{rcl@{\hspace{10mm}}l}
  \sty &::=& \sharety \mid \pubty \mid \outty & \textit{share types} \\
  \stt &::=& [\phi,\phi] \mid \pubx{\phi} & \textit{share terms} 
\end{array}
$$

\subsection{Non-Interactive Typing Rules}

\subsection{Interactive Typing Rules}

@@

@@

\subsection{Share Typing Example} To illustrate details of the
type system we return to the example of additive sharing in the context
of a 2-client protocol. We first focus on initial secret sharing steps--
let $\prog$ be the following protocol in $\fieldp{2}$ (noting
that $-$ is equivalent to xor in $\fieldp{2}$):
$$
\begin{array}{lll}
  \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip)}{1}; \\
  \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip)}{2} 
\end{array}
$$
We can then assert:
$$
\toeq{\prog} \vdash \varnothing,\varnothing \stredx
\setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}, \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
$$
Suppose the protocol goes on to reveal the negation of the
sum (xor-ing) of the inputs:
$$
\begin{array}{lll}
  \px{1} &:=& \elab{\mesg{s2} + \locflip + 1}{1}; \\
  \px{2} &:=& \elab{\mesg{s1} + \locflip}{2}\\
  \out{1} &:=& \elab{\px{1} + \px{2}}{1}\\
  \out{2} &:=& \elab{\px{1} + \px{2}}{2}
\end{array}
$$
Letting this extended protocol be $\prog'$, and letting:
$$
\begin{array}{rcl}
\Gamma &\defeq& \setit{[\mx{s2}{1}, \elab{\locflip}{2}] : \sharety,
  [\mx{s1}{2}, \elab{\locflip}{1}] : \sharety}\\
R &\defeq&  \setit{\elab{\locflip}{1},\elab{\locflip}{2}}
\end{array}
$$
observe we can then obtain share typings for noninteractive components for
addition of shares by \TirName{Shared} and \TirName{HeAdd}:
$$
\Gamma, \toeq{\prog'} \vdash
[\mx{s2}{1} + \elab{\locflip}{1}, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
$$
and hence for negation by \TirName{HeAddPub}:
$$
\Gamma, \toeq{\prog'} \vdash
[\mx{s2}{1} + \elab{\locflip}{1} + 1, \mx{s1}{2} + \elab{\locflip}{2}] : \sharety
$$
This implies:
$$
\toeq{\prog'} \vdash \varnothing,\varnothing \stredx
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, R 
$$
Hence by \TirName{PubEntails} we have:
$$
\begin{array}{l}
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
\pubx{\out{1}} : \outty \\
\Gamma \cup \setit{\pubx{\px{1} + \px{2}} : \outty}, \toeq{\prog'} \vdash
\pubx{\out{2}} : \outty
\end{array}
$$

@@

\begin{definition} Define $\ov(\Gamma)$ as follows:
\begin{eqnarray*}
\ov(\varnothing) &=& \varnothing \\[-1.5mm]
\ov(\setit{[\mx{z}{\cid}, \rx{y}{\cid'}] : \sharety} \cup \Gamma) &=& \setit{\mx{z}{\cid}} \cup \ov(\Gamma)\\[-1.5mm]
\ov(\setit{[\px{x} + \px{y}] : \pubty} \cup \Gamma) &=& \setit{\px{x},\px{y}} \cup \ov(\Gamma)\\[-1.5mm]
\ov(\setit{[\out{\cid}] : \outty} \cup \Gamma) &=& \setit{\out{\cid}} \cup \ov(\Gamma)
\end{eqnarray*}
\end{definition}
